---
title: "lab_02"
author: "derek willis"
date: "2024-08-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

For this lab we'll be working with a slice of Maryland voter registration data: records of people who have registration dates this year through Aug. 7. Let's load it.

```{r}
md_voters_2024 <- read_rds("data/lab_02.rds")
# Define the date of interest
date_cutoff <- as.Date("2024-07-21")

# Filter for new registrants after the cutoff date
new_voters <- md_voters_2024 |>
  filter(STATE_REGISTRATION_DATE == COUNTY_REGISTRATION_DATE, 
         STATE_REGISTRATION_DATE > date_cutoff)

# Count new voters by county
new_voters_by_county <- new_voters |>
  group_by(COUNTY) |>
  summarise(count = n())

# View the counties with the most new voters
new_voters_by_county
```

## Examine Data

Take a look at the data, getting familiar with what it contains. Most of the columns should be self-explanatory, but there are two date columns: STATE_REGISTRATION_DATE and COUNTY_REGISTRATION_DATE. The former represents the date the person registered as a Maryland voter; the latter is when they registered in their current county. They may be the same, but if a voter moves from one county to another they will differ. This is important because if the two dates are the same, that's a pretty good indication that this is a *new* voter, someone who has no history of registering in Maryland. You'll need to keep that definition in mind.

## Answer questions

**Q1** How many voters appear to be *new* registrants since July 21, when President Joe Biden announced he would leave the race, based on the information above? Create a dataframe to hold the only those voters. Which counties have seen the most? Take a look at county population figures here and compare the two in your answer: https://worldpopulationreview.com/us-counties/maryland.

**There are around 17,076 new registrants since the July 21 cutoff. Montgomery (4,847), and Baltimore City (2,799) were the counties that saw the most.**

```{r}
# Add week number to the dataframe
md_voters_2024 <- md_voters_2024 |>
  mutate(week_number = week(COUNTY_REGISTRATION_DATE))

# Count registrants by week
registrants_by_week <- md_voters_2024 |>
  group_by(week_number) |>
  summarise(count = n())

# View the distribution
registrants_by_week
```

**Q2** A better question is whether the new voters you found in Q1 are an outlier compared to other weeks this year. Using the original dataframe with 2024 registrants, use `mutate()` and lubridate to produce week numbers from the COUNTY_REGISTRATION_DATE (Lubridate has a function that turns a date into a week number. Use that to answer this question). Save that to the original dataframe. Then calculate how many voters are in each week. Describe the distribution of voters groups in a sentence - are the weeks from July 21-August 7 different from the rest, and if so, how?

**A2**

```{r}
# Add month information to the dataframe
md_voters_2024 <- md_voters_2024 |>
  mutate(month = month(COUNTY_REGISTRATION_DATE, label = TRUE))

# Count registrations by month
registrants_by_month <- md_voters_2024 |>
  group_by(month) |>
  summarise(count = n())

# View the month with the most registrations
registrants_by_month
```

**Q3** Using lubridate, show which month saw the largest number of registrations. Does the answer help shed more light on Q2?
**A3** 

```{r}
md_voters_2024 <- md_voters_2024 |>
  mutate(region = case_when(
    COUNTY %in% c("Baltimore City", "County1", "County2") ~ "Region1",
    COUNTY %in% c("County3", "County4") ~ "Region2",
    # Add all necessary mappings here
    TRUE ~ "Other"
  ))

# Count new voters by region
new_voters_by_region <- new_voters |>
  mutate(region = case_when(
    COUNTY %in% c("Baltimore City", "County1", "County2") ~ "Region1",
    COUNTY %in% c("County3", "County4") ~ "Region2",
    # Add all necessary mappings here
    TRUE ~ "Other"
  )) |>
  group_by(region) |>
  summarise(count = n())

# View the regions with the most new voters
new_voters_by_region

```

**Q4** Maryland has, officially, five regions: https://www.visitmaryland.org/article/maryland-regions. Modify the original dataframe again to add a `region` column based on the `COUNTY` value using case_when. Are you going to have to do this for every Maryland county plus Baltimore City? Yes, you are. Which regions and parties account for the most new voters?

**A4** 

```{r}
```

**Q5** What's an additional question you could ask of this data?

**What are the demographic characteristics of new registrants compared to existing registrants?Did major political events or announcements (besides the one mentioned) influence voter registration rates?** 
