---
title: "Reverse Engineering Project"
author: "Henry Lilienfeld"
date: "September 29, 2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, 
[In North Carolina, Black Voters’ Mail-In Ballots Much More Likely to Be Rejected Than Those From Any Other Race](https://www.propublica.org/article/in-north-carolina-black-voters-mail-in-ballots-much-more-likely-to-be-rejected-than-those-from-any-other-race)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(refinr)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

NewVoter_data <- read_csv("data/nonvoters_data.csv",
  guess_max = 2111797) |>
  clean_names() |>
  mutate(
    voter_zip = as.character(voter_zip),
    election_dt = as.Date(election_dt)
  )
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [“Black voters were more than twice as likely to have mail-in ballots rejected than those submitted by the state’s white voters in 2018, and rejection rates for 2020 show a similar pattern, according to a new analysis by ProPublica and WRAL News.” ]
* **Analysis summary**: [This paragraph establishes a foundation for the entire investigation. It is clear that original data analysis by the reporters unearthed racial disparities in ballot rejection rates. This finding is key to understanding the study's focus on systemic inequalities.]

```{r}
# Put code to reverse engineer sentence here
NewVoter_data |>
  group_by(race, ballot_rtn_status)|>
  summarise(count = n())|>
arrange(desc(count))



# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: [“Guilford County, the third-most populous county in North Carolina, rejected the largest number of mail-in ballots by Black voters statewide: 177. The rejection rate for Black voters, at 16.5%, was three times as high as the rejection rate for white voters, 5.5%, putting the county largely in line with the rest of the state.”]
* **Analysis summary**: [This specific example illustrates a racial disparity in one of the most populous counties, revealing how data analysis spotlighted local discrepancies and used them as evidence of broader statewide patterns.]

```{r}
# Put code to reverse engineer sentence here
NewVoter_data |>
  group_by(county_desc, race)|>
  summarise(count = n())|>
arrange(desc(count))

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [“So far, 2020 shows a similar pattern. As of Sept. 23, the rejection rate for mail-in ballots submitted by Black voters was about 3%, nearly three times as high as the rejection rate for white voters, according to data from the state.”]
* **Analysis summary**: [This paragraph shows the continuity of the trend, linking current data to the same issues observed in 2018. It confirms that disparities persist.]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-